---
title: "MNIST"
author: "Claudia Soto"
date: "27/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Well known examples for classification (with unsupervised learning)

The MNIST data set is a well known example for classification. Here, for academic porpuses we will use unsupervised learning to try to achieve this task. Let's do it by performing k-means clustering on the MNIST pixel features an see if we can identify unique clusters of digits without using the response variable. Here, we declare \(k = 10\) only because we already know there are 10 unique digits represented in the data.


```{r}
library(tidyverse)
library(dslabs)
library(tidymodels)

#install.packages("dslabs")
# https://bradleyboehmke.github.io/HOML/kmeans.html
mnist <- dslabs::read_mnist()

```

```{r}
glimpse(mnist)
#Las labels son el número que le corresponde

```
```{r}
#por el momento solo usamos images, y no labels
features<- mnist$train$images
glimpse(features)

#784 pixeles
```

But, I want to see the pictures, how? (with ggplot)
```{r}
str(features[1,])
d <- matrix(features[1,], nrow = 28, byrow = TRUE)
#matrix cuadrada 28x28

```


```{r}
library(reshape2)
#melt para pasar los datos a ggplot
# toma la matriz y los pone como var1 var2 value
#es como sistema de coordenas pixel (1,1) = value(0)
digito <- melt(d) 
#longData<-longData[longData$value!=0,]
digito
```


```{r}
theme_set(theme_light())
ggplot(digito, aes(x= Var2, y =Var1)) +
  geom_raster(aes(fill=value)) + scale_y_reverse()

#geom_raster para que imprima una imagen fill con el valor que se tiene en el pixele
```


```{r}
library(patchwork)
p <- function(i){
    d <- matrix(features[i,], nrow = 28, byrow = TRUE)
    digito<-melt(d)
    ggplot(digito, aes(x = Var2, y = Var1)) +
    geom_raster(aes(fill=value)) +  coord_fixed() +
      scale_y_reverse() +
      theme_void() + theme(legend.position = "none")
}


```
```{r}
p(1) + p(2) + p(3) + p(4) + p(5) + p(6)
```

Actividad: Hacer un análisis estadístico de la distribución de pixeles para cada tipo de dígito



```{r}
labels_dig <-  mnist$train$labels
labels_dig <- as.matrix(labels_dig)

features2  <- as_tibble(cbind(labels_dig, features))
```

```{r}
g <- function(i){
  
    contador <- features2 %>% 
      filter(V1 == i) %>% 
      select(-V1) %>% 
      melt(id.vars = NULL)
    
     ggplot(contador, aes(x = value)) +
       geom_histogram(binwidth = 5,fill = "darkslateblue") +
       theme(plot.title = element_text(color = "midnightblue", size = 8, face = "bold"), 
             axis.title = element_text(size=7))+
       labs(title = paste("Number:", i),x = 'Pixeles', y = 'Freq')
}

```

```{r}

g5 <- function(i){
  
    contador <- features2 %>% 
      filter(V1 == i) %>% 
      select(-V1) %>% 
      melt(id.vars = NULL) %>% 
      count(value, sort = TRUE) %>% 
      slice_max(n, n = 5)
    
     ggplot(contador, aes(x = factor(value), y = n)) +
       geom_bar(fill = "darkslateblue", stat = "identity") +
       theme(plot.title = element_text(color = "midnightblue", size = 8, face = "bold"),
             axis.title = element_text(size=7), axis.text.x = element_text(angle = 65, vjust = 0.7, size = 8)) +
       labs(title = paste("Top 5 Number:", i),x = 'Pixeles', y = 'Freq') 

}

```


```{r}
g(0) + g(1) + g(2) + g(3) + g(4) + g(5) + g(6) + g(7) +g(8) +g(9)    
```

```{r}
g5(0) + g5(1) + g5(2) + g5(3) + g5(4) + g5(5) + g5(6) + g5(7) +g5(8) +g5(9)   
```




```{r}
#Profe
labels_dig <-  mnist$train$labels


x <- features[1,] %>% as.data.frame()

for(i in 2:6000) {
  if(labels_dig[i] == 5){
    x <- rbind(x, features[i,] %>% as.data.frame())
  }
}



#as.data.frame(table(unlist(x)))

```


```{r}
ggplot(x,aes(x=.)) + geom_histogram(binwidth= 5)

```


```{r}
mnist_clustering <- kmeans(features, centers = 10, nstart = 10)
```


```{r}
str(mnist_clustering)
```


```{r}
# Extract cluster centers
mnist_centers <- mnist_clustering$centers
glimpse(mnist_centers)

```


```{r}

pk<- function(i){
    d <- matrix(mnist_centers[i,], nrow = 28, byrow = TRUE)
    digito<-melt(d)
    ggplot(digito, aes(x = Var2, y = Var1)) +
    geom_raster(aes(fill=value)) +  coord_fixed() +
      scale_y_reverse() +
      theme_void() + theme(legend.position = "none")
}

```


```{r}
pk(1) + pk(2) + pk(3) + pk(4) + pk(5) + pk(6) + pk(7) + pk(8) + pk(9) + pk(10)

```


```{r}
# Plot typical cluster digits
par(mfrow = c(2, 5), mar=c(0.5, 0.5, 0.5, 0.5))
layout(matrix(seq_len(nrow(mnist_centers)), 2, 5, byrow = FALSE))
for(i in seq_len(nrow(mnist_centers))) {
  image(matrix(mnist_centers[i, ], 28, 28)[, 28:1],xaxt="n", yaxt="n")
}

```


```{r}
# Create mode function
mode_fun <- function(x){  
  which.max(tabulate(x))
}

 

mnist_comparison <- data.frame(
  cluster = mnist_clustering$cluster,
  actual = mnist$train$labels
) %>%
  group_by(cluster) %>%
  mutate(mode = mode_fun(actual)) %>%
  ungroup() %>%
  mutate_all(factor, levels = 0:9)

 

# Create confusion matrix and plot results
yardstick::conf_mat(
  mnist_comparison, 
  truth = actual, 
  estimate = mode
) %>%
  autoplot(type = 'heatmap')
#si hay muchos valores fuera de la diagonal, las cosas no van tan bien, y te dice en dónde esta el problema
#identificar cuáles números fueron mal asignados 

```
